# Missing values

```{r, include=FALSE}
library(tidyverse)
library(patchwork)
library(janitor)
library(ggpubr)
library(egg)
```

```{r, include=FALSE}
plot_missing_pattern = function(mycars, show_percentage=FALSE){
# a tibble with each row: <var> <missing_count>, sorted desc
  if (show_percentage) {
    missing_patterns_by_column <- mycars %>%
      summarise_all(list(~ sum(is.na(.))/dim(mycars)[1])) %>%
      pivot_longer(names(mycars)) %>%
      arrange(desc(value))
    STRING1 = "% rows"
    
  } else {
    missing_patterns_by_column <- mycars %>%
      summarise_all(list(~ sum(is.na(.)))) %>%
      pivot_longer(names(mycars)) %>%
      arrange(desc(value))
    STRING1 = "num rows missing:"
  }

# save a copy of descend column names for future use
descend_column_names <- missing_patterns_by_column$name

column_missing_bar_chart <- ggplot(
  missing_patterns_by_column,
  aes(x=factor(name, levels=descend_column_names), y=value))+
  geom_bar(stat='identity', fill='#86A6EF') +
  ylab(STRING1) + xlab('') +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))

# the starting code, grouped by column, count missing, and add a total column
missing_patterns <- data.frame(is.na(mycars)) %>%
  group_by_all() %>%
  count(name = "count", sort = TRUE) %>%
  ungroup()

# find out the row with all FALSEs
missing_patterns_plot <- missing_patterns %>% mutate_all((as.integer)) %>% mutate(
  num.of.missing = rowSums(.[1:dim(mycars)[2]]), id=row.names(.)) %>%
  mutate(alpha=ifelse(num.of.missing==0, 1, 0.5))
complete_row <- missing_patterns_plot  %>% filter(num.of.missing == 0) %>% select(id) %>% as.character()

order_level <- sort(as.numeric(unique(missing_patterns_plot$id)))
missing_patterns_plot$id <- factor(missing_patterns_plot$id, levels = order_level)

row_missing_bar_chart <- missing_patterns_plot %>%
  ggplot(aes(y=fct_rev(id), x=count)) +
  geom_bar(stat='identity', fill='#86A6EF', aes(alpha=factor(alpha))) +
  xlab(STRING1) + ylab('') +
  theme(legend.position = "none") + 
  scale_alpha_manual(values = c("0.5"=0.5, "1"=1),guide="none") + 
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))


main_plot_df <- data.frame(is.na(mycars)) %>%
  group_by_all() %>%
  count(name = "count", sort = TRUE) %>%
  ungroup() %>%
  rownames_to_column("id") %>%
  select(-count) %>%
  gather(key, value, -id) %>%
  mutate(id = as.factor(id), missing = ifelse(value=="TRUE", "1","0"), key = factor(key, levels = c(descend_column_names)))

order_level <- sort(as.numeric(unique(main_plot_df$id)))
main_plot_df$id <- factor(main_plot_df$id, levels = order_level)

main_plot <- ggplot(main_plot_df) + geom_tile(aes(x = key, y = fct_rev(id), fill=missing), color = "white") + 
  scale_fill_manual(values=c("1"="#A689E0", "0"="white")) +
  annotate(geom="text", x=descend_column_names[round(dim(mycars)[2]/2)], y=complete_row, label="complete cases") +
  theme(legend.position="none") +
  xlab("variable") +
  ylab("Missing pattern") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust=1))


# a pure ggplot2 with white background
null.plot <- ggplot() + theme_minimal()
ggarrange(
  column_missing_bar_chart,
  null.plot,
  main_plot,
  row_missing_bar_chart,
  ncol=2,nrow=2,widths = c(20,5),heights=c(4,10),
  top="Missing value patterns"
)
}
```

```{r, include=FALSE}
df <- read.csv("C:/Users/20961/Desktop/2008.csv")
```

```{r}
plot_missing_pattern(df)
```

According to the first plot presented above, the columns/features CarrierDelay, WeatherDelay, NASDelay, SecurityDelay, and LateAircraftDelay have the most number of rows of missing values. These five reasons have missing values above 1,500,000 rows. On the other hand, the other 8 variables, including ArrTime, ActualElapsedTime, AirTime, etc.  some rows of missing values but the total rows of missing values are all less than 100,000.

There are six missing patterns in our dataset and the most common pattern shows that CarrierDelay, WeatherDelay, NASDelay, SecurityDelay, and LateAircraftDelay are missing altogether. Our assumption based on this observation is that most of the flights were not delayed or the reasons of those delays are unknown. The second most-common pattern is that all of the data are shown in the dataset, implying no missing data. The third-most-common pattern is that all of the features in the most-common pattern are missing, plus ArrTime, ActualElapsedTime, AirTime, ArrDelay, Taxiln, DepTime, DepDelay, and TaxiOut. Our assumption based on this observation is that those flights were canceled, so we have no related information about the delay, departure, arrival, etc. The other three patterns show that the number of missing values is much fewer than that of the three most-common features. Thus, we will not discuss them in detail here.